---
title: "Semester Project"
format:
  html:
    code-fold: true
    code-line-numbers: true
    code-link: true
editor: visual
output: html_document
execute:
  keep-md: true
---



```{r, message=FALSE}
library(tidyverse)
library(rio)
library(pander)
library(haven)
library(plotly)
library(readr)
library(foreign)
library(readxl)
library(ggplot2)
library(ggthemes)
library(DT) 
library(ggrepel)
library(car)
library(mosaic)


```

### Import 

```{r, message=FALSE, warning=FALSE}
Rent <- read_csv("../DS350/Rent1.csv")
```



### Data Interactive Chart

The interactive chart below shows you all the apartment complexes for BYUI housing. The + button displays details such as phone number and website. It also has a search bar so feel free to use it. 

<br />


```{r, na.rm = TRUE}

Rent1 <- Rent %>% drop_na()  

#fem_house <- Rent1 %>% 
             # filter(Gender == "F")

datatable(Rent1, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")
```
<br />

### Price & Capacity

The Capacity is greater for women reaching the most expensive rent but very close to men's housing. The median for men in capacity is 62 and for women 108

```{r, message=FALSE, warning=FALSE}


Rent1 %>% 
  filter(Capacity!= 0) %>% 
ggplot(aes(x= Capacity, y= Price, fill=Gender)) + 
  geom_boxplot()+
  geom_point()+
  facet_wrap(~ Gender)+
  labs()
  theme_clean()

```


```{r}

favstats(Capacity ~ Gender, data= Rent1)[,-10] %>%
pander(caption="Capacity")

```



```{r}
price.by.gender <- Rent1 %>% 
  group_by(Gender) %>% 
  summarize(avg.rent= mean(Price))
```


### Rent Price

Overall the average rent is close in price for both genders. We have to consider that this dataset is from 2018-2019 so prices have changed.

```{r}
ggplot(price.by.gender, aes(x= Gender, y= avg.rent, fill= Gender))+
  geom_col()+
  theme_clean()+
  labs(title="Average Rent by Gender")
```

```{r}

price.by.gender %>% pander()

```
### Distance

From the plot it can be said that the distance from campus does not determine the price. For example The Pines is less than 2 min. away from the MC and the rent is less than $100. While North Point less than 2 min away costs 1,537.
  
```{r, message=FALSE, warning=FALSE}
plot_ly(Rent1) %>%
  add_markers(y= ~Price, x= ~WalkingMinutes, text= ~Apartment, color= ~ Gender, size= 1) %>%
  layout(title="Housing Distance from Campus by Gender", xaxis=list(title="Distance from MC in Walking Minutes"), 
         yaxis=list(title="Rent Price for Semester"))
```

### Closest Apartments 

The Pines and Birch Plaza seem to be the best options when it comes to distance and price. 

```{r,na.rm = TRUE,  warning = FALSE, message = FALSE}


# Price and distance (5 min)

closest.ap<- Rent1 %>% 
  filter(WalkingMinutes <=5)

p <- plot_ly(closest.ap) %>%
  add_markers(y= ~Price, x= ~WalkingMinutes, text= ~Apartment, color= ~ Gender, size= 1) %>%
  layout(title="Get to Campus in 5 or Less Minutes Walking", xaxis=list(title="Distance from MC"), 
         yaxis=list(title="Rent Price per Semester"))
p
```



### Parking 

Hover over the plot X:Price, Y:# Parking Spots, Z: Gender

Nauvoo House and North Point have more than 300 parking stalls with a rent price of $1537 and 1279 respectively.


```{r}


parking <- plot_ly(Rent1) %>%
  add_markers(y= ~ ParkingStalls, x= ~ Price,text= ~Apartment, ~Gender, colors = c("#3399FF", "#FF6666"), size= 1.2) %>%
  layout(title="Paking Spots by Price", xaxis=list(title="Price"), 
         yaxis=list(title="Parking Stalls"))
parking
```

```{r}

favstats(ParkingStalls ~ Gender, data= Rent1)[,-10] %>%
pander(caption="Parking")

```
  
  <br />
  

### Interpretation 

Based on the above capacity, pricing, distance and parking evaluation, here are the apartments that would be the best options.
The cheapest and closest to campus apartments are Birch plaza and Carriage House. If the student has a car and likes to live in a big apartment complex the best options are North Point and Nauvoo House. 



<br />

<br />
 
[^1]: Source: https://rpubs.com/quincymccarney/852758

